---
title: "Combined Analysis of Loess Data"
author: "Julian Cooper"
date: "11/08/23"
output:
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

# Step 0: Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(agricolae)
library(tibble)
library(rmarkdown) # For formatting rmarkdown
library(minpack.lm) # for nlsLM function
library(cmstatr) # for calculating CV

```

# Step 1: Merge individual year loess files
All stand counts, weed filtering, and peak/valley parameters updated in the individual year code.
```{r}

# Load data from each year
years <- seq(2018,2021) # Get sequence of years
cc_all_env <- data.frame()

# Loop through each data frame and merge
for (year in years) {
  
  cc_env <- read.csv(paste0("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/", year,"/results/",year,"_loess_CC.csv"), check.names = FALSE)
  
  cc_all_env <- rbind(cc_all_env, cc_env)
}

# Plot growth curves and environmental means
# Takes a long time to run, keep commented out id doing work in chunk
cc_all_env %>%
  # Pivot data frame long
  pivot_longer(cols = 2:29, names_to = "GDD", values_to = "CC") %>%
  # Adjust structure
  mutate(GDD = as.numeric(GDD)) %>%
  mutate(Years = as.factor(Year)) %>%
  # Plot
  ggplot(aes(GDD, CC, color = Year)) +
  geom_line(aes(group=Plot), color="gray") +
  stat_smooth(method="loess", formula = y ~ x) +
  stat_summary(aes(GDD, CC, group = Year, color = Year), geom = 'point', fun = mean, shape = 17, size = 3) +
  xlab('Growing Degree Day (GDD)') +
  ylab('Canopy Cover %')

# Save data frame
write.csv(cc_all_env, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/1loess_combo.csv", row.names = FALSE)

# Pivot long
cc_all_env_long <- cc_all_env %>%  
  pivot_longer(cols = 2:29, names_to = "GDD", values_to = "CC")
  
# Save long data frame
write.csv(cc_all_env_long, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/2loess_combo_long.csv", row.names = FALSE)

############ Stats for % CC ##################
# Load data
cc_all_env_long <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/2loess_combo_long.csv")

cc_all_env <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/1loess_combo.csv")

# EXPERIMENTAL DESIGN
# Count unique genotypes
cc_all_env_long %>%
    summarise(n_distinct(Genotype)) #499

# Count unique genotypes per year
cc_all_env_long %>%
  group_by(Year) %>%
  summarise(n_distinct(Genotype)) %>%
  ungroup()

# Count plots per year
cc_all_env %>%
  group_by(Year) %>%
  summarise(n=n()) %>%
  ungroup()

# Average reps per genotype
cc_all_env %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count < 9) %>% #Filter out PH207, B73, and ND259 check lines
  summarise(mean(rep_count))

# Tukey T-test for Year
## Check structure
cc_all_env$Rep <- as.factor(cc_all_env$Rep)
cc_all_env$Block <- as.factor(cc_all_env$Block)
cc_all_env$Genotype <- as.factor(cc_all_env$Genotype)
cc_all_env$Year <- as.factor(cc_all_env$Year)

## Check statistical differences between years for 1450 GDD, highest shared terminal GDD 
mod <- aov(cc_all_env[,25] ~ Year, data = cc_all_env)
mod_means_contr <- emmeans::emmeans(object = mod,
                                    pairwise ~ "Year",
                                    adjust = "tukey")
multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)

# PERCENT CANOPY COVER
# Average, Max, Min, and CV % CC of each GDD
sup_fig_3 <- cc_all_env_long %>%
  group_by(GDD) %>%
  mutate(CC = CC * 100) %>%
  summarise(meanCC = round(mean(CC, na.rm = TRUE), 3),
            maxCC = round(max(CC, na.rm = TRUE), 3),
            minCC = round(min(CC, na.rm = TRUE),3),
            CV = round(cv(CC, na.rm = TRUE), 3),
            SD = round(sd(CC, na.rm = TRUE), 3))

# Average, Max, Min, and CV % CC of each GDD
x <- cc_all_env_long %>%
  mutate(Stage = case_when(GDD %in% c(seq(300,450)) ~ "Juv.",
                           GDD %in% c(seq(500,1100)) ~ "Adult",
                           GDD %in% c(seq(1150,1650)) ~"Rep.")) %>%
  group_by(Stage) %>%
  mutate(CC = CC * 100) %>%
  summarise(mean = round(mean(CC, na.rm = TRUE), 3),
            max = round(max(CC, na.rm = TRUE), 3),
            min = round(min(CC, na.rm = TRUE),3),
            CV = round(cv(CC, na.rm = TRUE), 3),
            SD = round(sd(CC, na.rm = TRUE), 3))

# Tukey T-test for Year 
## Check structure
cc_all_env$Rep <- as.factor(cc_all_env$Rep)
cc_all_env$Block <- as.factor(cc_all_env$Block)
cc_all_env$Genotype <- as.factor(cc_all_env$Genotype)
cc_all_env$Year <- as.factor(cc_all_env$Year)

## Check statistical differences between years for ALL GDD
# Create an empty data frame to store results
tukey_results_df <- data.frame()

# Loop through values of I
for (i in 2:29) {
  # Fit the model
  mod <- aov(cc_all_env[, i] ~ Year, data = cc_all_env)
  
  # Calculate Tukey groups
  mod_means_contr <- emmeans::emmeans(
    object = mod,
    pairwise ~ "Year",
    adjust = "tukey"
  )
  
  # Get Tukey groups and convert to a data frame
  tukey_groups <- as.data.frame(multcomp::cld(
    object = mod_means_contr$emmeans,
    Letters = letters
  ))
  
  # Add a column to indicate the variable (I value)
  tukey_groups$I <- i
  
  # Append the results to the main data frame
  tukey_results_df <- rbind(tukey_results_df, tukey_groups)
}

# Print or view the final data frame
print(tukey_results_df)

```

# Step 2: Area Under Progress Curve (AUPC)
Not used in paper
```{r}

# Load data
c_full <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/1loess_combo.csv", check.names = FALSE)

# Only use flights starting at 450 and ending 1450 to keep AUPC even between years
c <- c_full[,c(1, 5:25, 30:34)]

# Make list of column names to extract dates
dates <- colnames(c)

# Trim list
dates <- dates[2:22]

# Change to numeric
dates <- as.numeric(dates)

# Specify which columns to evaluate
evaluation <- c[,2:22]

# Calculate AUPC
AUPC <-audpc(evaluation, dates)

# Change to data frame
AUPC <- tibble(AUPC)

# Add to "c" data frame
c_AUPC <- cbind(c_full, AUPC)

# Plot histogram of AUPC
c_AUPC %>%
  ggplot(aes(AUPC)) +
  geom_histogram() +
  facet_grid(~Year) # Facet by Env.

# Save data
write.csv(c_AUPC, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/3loess_combo_AUPC.csv", row.names = FALSE)

############ Stats for AUPC ##################

# Mean and SD AUPC
c_AUPC %>%
 summarise(mean = round(mean(AUPC, na.rm = TRUE), 3),
            max = round(max(AUPC, na.rm = TRUE), 3),
            min = round(min(AUPC, na.rm = TRUE),3),
            sd = round(sd(AUPC, na.rm = TRUE), 3),
            CV = round(cv(AUPC, na.rm = TRUE), 3))

# Mean and SD AUPC per year
c_AUPC %>%
  group_by(Year) %>%
  summarise(mean(AUPC, na.rm = TRUE),
            sd(AUPC, na.rm = TRUE)) %>%
  ungroup()

```

# Step 3: Rate of canopy development between 50 GDD intervals
```{r slope}

# Load data 
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/3loess_combo_AUPC.csv", check.names = FALSE)

# Make list of column names to extract dates
dates <- as.numeric(colnames(c)[2:29])

# Calculate slope between each GDD
## Make empty data frame with 27 rows to output results
slope_data <- data.frame(matrix(NA, nrow = 27))

for (i in seq_len(nrow(c))) { #For each row in data frame
  row <- c[i,2:29] #Subset row
  
  # Slope formula y2-y1/x2-x2
  # Set sliding window for pulling slope numbers
  x1 <- 1
  x2 <- 2
  y2 <- 2
  
  # Make new data frame to save slopes for each row in
  slope_row <- data.frame()

  # For each column in extracted row (1-28) except for last (28), because no slope without a next point
  for (j in 1:27) {

  slope <- (((row[y2] - row[j]) / (dates[x2] - dates[x1]))) * 100 # Slope formula, multiplied by 100
  
  slope_row[j, 1] <- slope # Add to row data frame
  
  # Move to next time and data points to calculate new slope
  y2 <- y2 + 1
  x2 <- x2 + 1
  x1 <- x1 + 1
  }
  
  # Bind all slopes from each row as a new column in master data frame
  slope_data <- cbind(slope_data, slope_row)
}

## Format results
# Get list of plot names
plots <- c(c$Plot)

# Remove filler column
slope_data <- slope_data[, -1]

# Rename slope columns as corresponding plots
colnames(slope_data) <- plots

# Make new columns with slope order
slope_data$Slope <- seq(1:27)

# Transpose
slope_data_transpose <- t(slope_data)

# Change to data frame
slope_data_transpose <- as.data.frame(slope_data_transpose)

# Make row index (Plot) a new column
slope_data_transpose <- tibble::rownames_to_column(slope_data_transpose, "Plot")

# Merge slope with CC and AUPC
c_AUPC_slope <- merge(c, slope_data_transpose, by = "Plot")

# Save data frame
write.csv(c_AUPC_slope, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/4loess_combo_AUPC_slope.csv", row.names = FALSE)

# Plot distribution of slopes over time
## Subset df, only keep slope
c_slope <- c_AUPC_slope[, c(1, 30:34, 36:62)]

## Pivot long
c_slope_long <- pivot_longer(c_slope, cols = 7:33, names_to = "Interval", values_to = "Slope")

## Set order of Interval factor
c_slope_long$Interval = factor(c_slope_long$Interval, levels = c(seq(1:27)))

## Plot
ggplot(c_slope_long, aes(x=as.factor(Interval), y=Slope)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("GDD Interval")

############ Stats for Rate of CC Development ##################
# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/4loess_combo_AUPC_slope.csv", check.names = FALSE)

## Subset df, only keep slope
c_slope <- c[, c(1, 30:34, 36:62)]

## Pivot long
c_slope_long <- pivot_longer(c_slope, cols = 7:33, names_to = "Interval", values_to = "Slope")

# Convert intervals to GDD
c_slope_long <- c_slope_long %>%
  mutate(
    GDD = case_when(
      Interval == "1" ~ "300-350",
      Interval == "2" ~ "350-400",
      Interval == "3" ~ "400-450",
      Interval == "4" ~ "450-500",
      Interval == "5" ~ "500-550",
      Interval == "6" ~ "550-600",
      Interval == "7" ~ "600-650",
      Interval == "8" ~ "650-700",
      Interval == "9" ~ "700-750",
      Interval == "10" ~ "750-800",
      Interval == "11" ~ "800-850",
      Interval == "12" ~ "850-900",
      Interval == "13" ~ "900-950",
      Interval == "14" ~ "950-1000",
      Interval == "15" ~ "1000-1050",
      Interval == "16" ~ "1050-1100",
      Interval == "17" ~ "1100-1150",
      Interval == "18" ~ "1150-1200",
      Interval == "19" ~ "1200-1250",
      Interval == "20" ~ "1250-1300",
      Interval == "21" ~ "1300-1350",
      Interval == "22" ~ "1350-1400",
      Interval == "23" ~ "1400-1450",
      Interval == "24" ~ "1450-1500",
      Interval == "25" ~ "1500-1550",
      Interval == "26" ~ "1550-1600",
      Interval == "27" ~ "1600-1650")
  )

# Set order
GDD_order <- as.character(c("300-350", "350-400", "400-450", "450-500", "500-550", "550-600", "600-650", "650-700", "700-750",  "750-800",  "800-850",  "850-900",  "900-950",  "950-1000",  "1000-1050", "1050-1100",  "1100-1150",  "1150-1200",  "1200-1250",  "1250-1300",  "1300-1350",  "1350-1400",  "1400-1450",  "1450-1500",  "1500-1550",  "1550-1600",  "1600-1650"))

# Average, Max, Min, and SD Slope of each GDD
rate_stats <- c_slope_long %>%
  group_by(GDD) %>%
  summarise(meanSlope = round(mean(Slope, na.rm = TRUE), 2),
            maxSlope = round(max(Slope, na.rm = TRUE), 2),
            minSlope = round(min(Slope, na.rm = TRUE),2),
            sdSlope = round(sd(Slope, na.rm = TRUE), 2),
            CV = round(cv(Slope, na.rm = TRUE), 3)) %>%
  arrange(factor(GDD, levels = GDD_order))

# Max Slope of each Year
c_slope_long %>%
  group_by(Year) %>%
  filter(Slope == max(Slope, na.rm = TRUE))

# Add Phase
juvenile <- c("300-350", "350-400", "400-450")
adult <- c("450-500", "500-550", "550-600", "600-650", "650-700", "700-750",  "750-800",  "800-850",  "850-900",  "900-950",  "950-1000",  "1000-1050", "1050-1100",  "1100-1150")
reproductive <- c("1150-1200",  "1200-1250",  "1250-1300",  "1300-1350",  "1350-1400",  "1400-1450",  "1450-1500",  "1500-1550",  "1550-1600",  "1600-1650")

# Average, Max, Min, and SD Slope of each Growth Stage
x <- c_slope_long %>%
  mutate(Stage = case_when(GDD %in% juvenile ~ "Juvenile",
                           GDD %in% adult ~ "Adult",
                           GDD %in% reproductive ~"Reproductive")) %>%
  group_by(Stage) %>%
   summarise(meanSlope = round(mean(Slope, na.rm = TRUE), 2),
            maxSlope = round(max(Slope, na.rm = TRUE), 2),
            minSlope = round(min(Slope, na.rm = TRUE),2),
            sdSlope = round(sd(Slope, na.rm = TRUE), 2),
            CV = round(cv(Slope, na.rm = TRUE), 3))

# Max rate of development for each growth stage in each year
paged_table(c_slope_long %>%
  mutate(Stage = case_when(GDD %in% juvenile ~ "Juvenile",
                           GDD %in% adult ~ "Adult",
                           GDD %in% reproductive ~"Reproductive")) %>%
  group_by(Year, Stage) %>%
  summarise(meanSlope = round(mean(Slope, na.rm = TRUE), 2),
            maxSlope = round(max(Slope, na.rm = TRUE), 2),
            minSlope = round(min(Slope, na.rm = TRUE),2),
            sdSlope = round(sd(Slope, na.rm = TRUE), 2)))

```

# Step 4: Adult Vegetative Slope
Not used in paper
```{r AVS}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/data/data_processing/4loess_combo_AUPC_slope.csv", check.names = FALSE)

# Calculate slope between 600 and 1000 GDD
c_AVS <- c %>%
    mutate(AVS = ((`1000` - `600`)/400)*100) # Slope formula, multiplied by 100

# Save final data frame with all traits results folder
write.csv(c_AVS, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", row.names = FALSE)

# Plot histogram of AVS
c_AVS %>%
  ggplot(aes(AVS)) +
  geom_histogram() +
  facet_grid(~Year) # Facet by Env.

############ Stats for AUPC ##################

# Mean and SD AVS
c_AVS %>%
 summarise(mean = round(mean(AVS, na.rm = TRUE), 3),
            max = round(max(AVS, na.rm = TRUE), 3),
            min = round(min(AVS, na.rm = TRUE),3),
            sd = round(sd(AVS, na.rm = TRUE), 3),
           CV = round(cv(AVS, na.rm = TRUE), 3))

# Mean and SD AVS per year
c_AVS %>%
  group_by(Year) %>%
  summarise(mean(AVS, na.rm = TRUE),
            sd(AVS, na.rm = TRUE)) %>%
  ungroup()

```

# Why not use the loess regression coefficients as a phenotypic parameter?
"The local polynomials fit to each subset of the data are almost always of first or second degree; that is, either locally linear (in the straight line sense) or locally quadratic. Using a zero degree polynomial turns LOESS into a weighted moving average. Higher-degree polynomials would work in theory, but yield models that are not really in the spirit of LOESS. LOESS is based on the ideas that any function can be well approximated in a small neighborhood by a low-order polynomial and that simple models can be fit to data easily. High-degree polynomials would tend to overfit the data in each subset and are numerically unstable, making accurate computations difficult."

"As discussed above, the biggest advantage LOESS has over many other methods is the process of fitting a model to the sample data does not begin with the specification of a function. Instead the analyst only has to provide a smoothing parameter value and the degree of the local polynomial. In addition, LOESS is very flexible, making it ideal for modeling complex processes for which no theoretical models exist."

"Another disadvantage of LOESS is the fact that it does not produce a regression function that is easily represented by a mathematical formula. This can make it difficult to transfer the results of an analysis to other people. In order to transfer the regression function to another person, they would need the data set and software for LOESS calculations."

TLDR; There are no uniform model parameters from Loess to use as phenotypes. Each data point is individually fit based on the span. This leads to a very accurate growth curve across the season, but at the cost of having a uniform model each plot is fit into. 
