---
Yestitle: "Dynamic Stability: FW Slope"
author: "Julian Cooper"
date: "11/13/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Background

## Bayesian Regression
Bayesian statistics is a mathematical procedure that applies probabilities to statistical problems to represent uncertainty. It provides researchers the tools to update their beliefs in the evidence of new data.

Example: I have misplaced my phone somewhere in the home. I can use the phone locator on the base of the instrument to locate the phone and when I press the phone locator the phone starts beeping. Which area of my home should I search?

Requires data, a generative model, and prior (info the model has before seeing the data).

Generative model: Feed model fixed parameters values and get simulated data in return. Useful for seeing how much data varies (Monte Carlo simulation). Can also be flipped. If you know the data, you can feed it into a generative model to find the unknown parameters.

Prior: Probability to represent uncertainty in the model. 

If you know the data, you can simulate parameters based on prior distributions until your generative model matches your observed data. For example, If 16 people see an add and 6 buy the product, you can assume the the probability of buying is 0-1.0, and run 100,000s of iterations of your generative model creating simulated data until you find a statistically significant parameter (probability) that matches your actual purchase rate of 6/16.

Posterior: Distribution (probability) of the parameter. 

Keeping with our advertisement example, if the prior distribution is uniform 0-1.0, after many iterations we see that our actual posterior parameter distribution is a normal curve centered between .3 and .4.

Posterior distribution is the end value from our data. Maximum likelihood estimate for our parameter is the highest probability parameter value from our posterior distribution. Could also use posterior mean of distribution. Often, a 90% credible interval is given covering the most likely parameter values. 

## Markov Chain Monte Carlo (MCMC)
A class of algorithms for sampling from a probability distribution. By constructing a Markov chain that has the desired distribution as its equilibrium distribution, one can obtain a sample of the desired distribution by recording states from the chain. The more steps are included, the more closely the distribution of the sample matches the actual desired distribution. This is how the posterior distribution is obtained from the prior in Bayesian regression.

## Gibbs Sampler
An MCMC algorithm for sampling for a multivariate distribution when sampling from joint distribution (getting a pair of x and y values) is difficult but conditional distibution sampling (x when y or y when x) is simple. Procedure: 1) Initialize some x,y value from data. 2) Keep y fixed, change new x. 3) Sample new y given fixed x from step 2. 4) Repeat until all data points have been sampled. 

## Finley Wilkinson Regression
Assesses how expected G performance varies as a function of the environment. Achieved by regressing the performance of each genotype on the environment mean.

# Step 0: Install Package
```{r setup}

library(devtools)
install_github("lian0090/FW")
library(FW)
library(coda)
library(tidyverse)
library(ggplot2)
library(MASS)

```

## Step 0.1: Test package
http://nitro.biosci.arizona.edu/zbook/newvolume_2/pdf/Chapter44.pdf  
https://rpubs.com/alice_macqueen/Field_FW_GxE_Analysis_2018  
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4777122/#box1  
https://www.semanticscholar.org/paper/FW%3A-An-R-Package-for-Finlay–Wilkinson-Regression-Lian-Campos/1510be011253057cbd80476a0edbee70b503df9b

```{r test_data}

# Load test data from FW package
data(wheat)

#The database is attached to the R search path. This means that the database is searched by R when evaluating a variable, so objects in the database can be accessed by simply giving their names.
attach(wheat.Y)

# Ordinary Least Squares Method
OLS <- FW(y = y,VAR = VAR,ENV = ENV,method="OLS")

# Bayesian FW regression assuming independence of lines and of environments
GibbsI = FW(y = y,VAR = VAR,ENV = ENV, method="Gibbs",seed = 12345,saveAt="GibbsI",nIter =
50000 ,burnIn = 5000)

# Bayesian FW regression that incorporates genomic information 
GibbsA = FW(y = y,VAR = VAR,ENV = ENV, method="Gibbs",A = wheat.G,seed = 12345, saveAt="GibbsA",nIter = 50000,burnIn = 5000) 

# The fitness of the models can be examined by the correlations between the observed values and the fitted values 
cor(y,OLS$yhat) 
cor(y,GibbsI$yhat)
cor(y,GibbsA$yhat)

# The correlations among parameter estimates from different models
cor(OLS$b,GibbsI$b)
cor(OLS$b,GibbsA$b)
cor(GibbsI$b,GibbsA$b)

# Plot of variety performance vs. estimated environment values. Each line represents a different variety. Lines are fitted values and points are the cell means of genotype and environment combination. The horizontal axis displays the estimated environmental effects. The labels of these environments are also displayed; these labels can be removed by setting ENVlabel = F.

plot.FW(OLS, main="OLS", cex = 0.2,lwd = 0.2)
plot.FW(GibbsA,main="GibbsA", cex = 0.2,lwd = 0.2)

```


# Step 1: Finley Wilkinson on AUPC
Not included in paper
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset into AUPC data only
c_AUPC <- c[, c(1, 30:35)]

# Filter out genotypes with rep count < 5
c_AUPC <- c_AUPC %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 

# Convert back to data frame
c_AUPC <- as.data.frame(c_AUPC)

# Remove NA
c_area <- na.omit(c_AUPC)

# Make empty date frame
area_FW_b <- data.frame()
area_FW_yhat <- data.frame()

# For each column with point data
for (i in 7:7){
  
  print(i)
  
  # Set model
  GibbsI = FW(y = c_area[,i],VAR = c_area$Genotype,ENV = c_area$Year, method="Gibbs",seed=12345,saveAt="GibbsI",nIter = 50000 ,burnIn = 5000)
  
  # Genotype Slope
  b <- data.frame(GibbsI$b)
  colnames(b) <- "b"

  # Predicted y values
  yhat <- data.frame(GibbsI$yhat)
  colnames(yhat) <- "yhat"

  # Observed y values
  y <- data.frame(GibbsI$y)
  colnames(y) <- "y"
  
  # Genotype
  g <- c_area$Genotype
  
  # Year
  year <- c_area$Year
  
  # Add column with data date to both data frames
  GDD <- colnames(c_area[i]) # Save as object
  b$GDD <- GDD
  
  yhat$year <- year
  yhat$Genotype <- g
  yhat$y <- y
  yhat$GDD <- GDD
  
  # Keep saving to original empty data frame
  area_FW_b <- rbind(area_FW_b, b)  
  area_FW_yhat <- rbind(area_FW_yhat, yhat)
}

# Format slope data
# Change Genotype to index to own column
area_FW_b  <- tibble::rownames_to_column(area_FW_b, "Genotype")

# Save results
write.csv(area_FW_b, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_AUPC_b.csv", row.names = FALSE)

# Format yhat data
y_yhat <- area_FW_yhat

#Unnest y list from data frame
y_yhat <- unnest(y_yhat, cols = 4)

# Save results
write.csv(y_yhat, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_AUPC_y.csv", row.names = FALSE)

## AUPC FW Stats
AUPC_FW <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_AUPC_b.csv")

stats <- AUPC_FW %>%
  mutate(b = b + 1) %>%
  summarise(meanFW = round(mean(b, na.rm = TRUE), 3),
            SDFW = round(sd(b, na.rm = TRUE), 3),
            CVFW = round(cv(b, na.rm = TRUE), 3),
            maxFW = round(max(b), 3),
            minFW = round(min(b, na.rm = TRUE), 3))
```

# Step 2: Finley Wilkinson on Point
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset into point
c_point_total <- c[, 1:34]

# Make empty date frame
point_FW_b <- data.frame()
point_FW_yhat <- data.frame()

# For each column with point data
for (i in 2:29){ #2:29
  
   # Make new data farme
  c_point <- c_point_total[,c(1, i, 30:34)]
  
  # Remove missing
  c_point <- na.omit(c_point)
  
  # Filter genotypes with less than 5 reps
  c_point <- c_point %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 
  
  # Make sure data is in data frame form
  c_point <- as.data.frame(c_point)
  
  print(i)
  
  # Set model
  GibbsI = FW(y = c_point[,2],VAR = c_point$Genotype,ENV = c_point$Year, method="Gibbs",seed=12345,saveAt="GibbsI",nIter = 50000 ,burnIn = 5000)
  
  # Genotype Slope
  b <- data.frame(GibbsI$b)
  colnames(b) <- "b"
  
  b <- tibble::rownames_to_column(b, "Genotype")

  # Predicted y values
  yhat <- data.frame(GibbsI$yhat)
  colnames(yhat) <- "yhat"

  # Observed y values
  y <- data.frame(GibbsI$y)
  colnames(y) <- "y"
  
  # Genotype
  g <- c_point$Genotype
  
  # Year
  year <- c_point$Year
  
  # Add column with data date to both data frames
  
  GDD <- colnames(c_point[2]) # Save as object
  b$GDD <- GDD

  yhat$year <- year
  yhat$Genotype <- g
  yhat$y <- y
  yhat$GDD <- GDD
  
  # Keep saving to original empty data frame
  point_FW_b <- rbind(point_FW_b, b)  
  point_FW_yhat <- rbind(point_FW_yhat, yhat)
  
}

# Give unique indentifiers and pivot wide
b_wide <- point_FW_b %>%
  tidyr::pivot_wider(names_from = GDD, values_from = b) 

# Save results
write.csv(b_wide, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_point_b.csv", row.names = FALSE)

# Format yhat data
y_yhat <- point_FW_yhat

#Unnest y list from data frame
y_yhat <- unnest(y_yhat, cols = 4)

# Save results
write.csv(y_yhat, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_point_y.csv", row.names = FALSE)

############ Stats for % CC ##################
point_FW <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_point_b.csv", check.names = FALSE)

# Pivot long
point_FW_longer <- point_FW %>%
  pivot_longer(cols = c(2:29), names_to = "GDD", values_to = "FW")

# Change GDD to numeric
point_FW_longer$GDD <- as.numeric(point_FW_longer$GDD)

## Average FW at each GDD
supf4 <- point_FW_longer %>%
  group_by(GDD) %>%
  mutate(FW = FW + 1)  %>%
  summarise(meanFW = round(mean(FW, na.rm = TRUE), 3),
            SDFW = round(sd(FW, na.rm = TRUE), 3),
            CVFW = round(cv(FW, na.rm = TRUE), 3),
            maxFW = round(max(FW, na.rm = TRUE), 3),
            minFW = round(min(FW, na.rm = TRUE), 3))

## Average FW at each Growth Stage
point_FW_longer %>%
  mutate(Stage = case_when(GDD %in% c(seq(300,450)) ~ "Juv.",
                           GDD %in% c(seq(500,1100)) ~ "Adult",
                           GDD %in% c(seq(1150,1650)) ~"Rep.")) %>%
  group_by(Stage) %>%
  mutate(FW = FW + 1) %>%
  summarise(meanFW = round(mean(FW, na.rm = TRUE), 3),
            SDFW = round(sd(FW, na.rm = TRUE), 3),
            CVFW = round(cv(FW, na.rm = TRUE), 3),
            maxFW = round(max(FW, na.rm = TRUE), 3),
            minFW = round(min(FW, na.rm = TRUE), 3))

```

# Step 3: Finley Wilkinson on Slope
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset into slope
c_slope_total <- c[, c(1, 30:34, 36:62)]

# Make empty date frame
slope_FW_b <- data.frame()
slope_FW_yhat <- data.frame()

# For each column with slope data
for (i in 7:33){ #2:29
  
   # Make new data farme
  c_slope <- c_slope_total[,c(1:6, i)]
  
  # Remove missing
  c_slope <- na.omit(c_slope)
  
  # Filter genotypes with less than 5 reps
  c_slope <- c_slope %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 
  
  # Make sure data is in data frame form
  c_slope <- as.data.frame(c_slope)
  
  print(i)
  
  # Set model
  GibbsI = FW(y = c_slope[,7],VAR = c_slope$Genotype,ENV = c_slope$Year, method="Gibbs",seed=12345,saveAt="GibbsI",nIter = 50000 ,burnIn = 5000)
  
  # Genotype Slope
  b <- data.frame(GibbsI$b)
  colnames(b) <- "b"
  
  b <- tibble::rownames_to_column(b, "Genotype")

  # Predicted y values
  yhat <- data.frame(GibbsI$yhat)
  colnames(yhat) <- "yhat"

  # Observed y values
  y <- data.frame(GibbsI$y)
  colnames(y) <- "y"
  
  # Genotype
  g <- c_slope$Genotype
  
  # Year
  year <- c_slope$Year
  
  # Add column with data date to both data frames
  
  GDD <- colnames(c_slope[7]) # Save as object
  b$GDD <- GDD

  yhat$year <- year
  yhat$Genotype <- g
  yhat$y <- y
  yhat$GDD <- GDD
  
  # Keep saving to original empty data frame
  slope_FW_b <- rbind(slope_FW_b, b)  
  slope_FW_yhat <- rbind(slope_FW_yhat, yhat)
  
}

# Give unique indentifiers and pivot wide
b_wide <- slope_FW_b %>%
  tidyr::pivot_wider(names_from = GDD, values_from = b) 

# Save results
write.csv(b_wide, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_slope_b.csv", row.names = FALSE)

# Format yhat data
y_yhat <- slope_FW_yhat

#Unnest y list from data frame
y_yhat <- unnest(y_yhat, cols = 4)

# Save results
write.csv(y_yhat, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_slope_y.csv", row.names = FALSE)

##### Slope FW Stats #######
slope_FW <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_slope_b.csv", check.names = FALSE)

# For each GDD
stats <- slope_FW %>%
  pivot_longer(cols = 2:28, names_to = "GDD", values_to = "FW") %>%
  mutate(FW = FW +1,
         GDD = as.numeric(GDD)) %>%
  group_by(GDD) %>%
summarise(meanFW = round(mean(FW, na.rm = TRUE), 3),
            SDFW = round(sd(FW, na.rm = TRUE), 3),
            CVFW = round(cv(FW, na.rm = TRUE), 3),
            maxFW = round(max(FW, na.rm = TRUE), 3),
            minFW = round(min(FW, na.rm = TRUE), 3))

#For each growth stage
# Add Phase
juvenile <- c("300-350", "350-400", "400-450")
adult <- c("450-500", "500-550", "550-600", "600-650", "650-700", "700-750",  "750-800",  "800-850",  "850-900",  "900-950",  "950-1000",  "1000-1050", "1050-1100",  "1100-1150")
reproductive <- c("1150-1200",  "1200-1250",  "1250-1300",  "1300-1350",  "1350-1400",  "1400-1450",  "1450-1500",  "1500-1550",  "1550-1600",  "1600-1650")

# Average, Max, Min, and SD Slope of each Growth Stage
stats <- slope_FW %>%
  pivot_longer(cols = 2:28, names_to = "GDD", values_to = "FW") %>%
  mutate(FW = FW +1,
    GDD = case_when(
      GDD == "1" ~ "300-350",
      GDD == "2" ~ "350-400",
      GDD == "3" ~ "400-450",
      GDD == "4" ~ "450-500",
      GDD == "5" ~ "500-550",
      GDD == "6" ~ "550-600",
      GDD == "7" ~ "600-650",
      GDD == "8" ~ "650-700",
      GDD == "9" ~ "700-750",
      GDD == "10" ~ "750-800",
      GDD == "11" ~ "800-850",
      GDD == "12" ~ "850-900",
      GDD == "13" ~ "900-950",
      GDD == "14" ~ "950-1000",
      GDD == "15" ~ "1000-1050",
      GDD == "16" ~ "1050-1100",
      GDD == "17" ~ "1100-1150",
      GDD == "18" ~ "1150-1200",
      GDD == "19" ~ "1200-1250",
      GDD == "20" ~ "1250-1300",
      GDD == "21" ~ "1300-1350",
      GDD == "22" ~ "1350-1400",
      GDD == "23" ~ "1400-1450",
      GDD == "24" ~ "1450-1500",
      GDD == "25" ~ "1500-1550",
      GDD == "26" ~ "1550-1600",
      GDD == "27" ~ "1600-1650")
  ) %>% 
  mutate(Stage = case_when(GDD %in% juvenile ~ "Juvenile",
                           GDD %in% adult ~ "Adult",
                           GDD %in% reproductive ~"Reproductive")) %>%
  group_by(Stage) %>%
summarise(meanFW = round(mean(FW, na.rm = TRUE), 3),
            SDFW = round(sd(FW, na.rm = TRUE), 3),
            CVFW = round(cv(FW, na.rm = TRUE), 3),
            maxFW = round(max(FW, na.rm = TRUE), 3),
            minFW = round(min(FW, na.rm = TRUE), 3))

```

## Step 4: Finley Wilkinson on AVS
Not included in paper
```{r eval=FALSE}
# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset into AUPC data only
c_expslope <- c[, c(1, 30:34, 63)]

# Filter out genotypes with rep count < 5
c_expslope <- c_expslope %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 

c_expslope %>%
  summarise(n_distinct(Genotype))

# Code testing if scale of original trait changes range of slope values
# Multiplying exp slope by 10000 changes FW slopes to range from -.8 - .8 but does not change relative order of genotypes (top and bottom stay ~about the same, some small variation in order but not by a lot)

# 10/13 multiple by 100 (not done in initial step during loess, to see if that affects FW scale)
c_expslope$Exp <- c_expslope$Exp * 100

# Convert back to data frame
c_expslope <- as.data.frame(c_expslope)

# Remove NA
c_expslope <- na.omit(c_expslope)

# Make empty date frame
expslope_FW_b <- data.frame()
expslope_FW_yhat <- data.frame()

# For each column with point data
for (i in 7:7){
  
  print(i)
  
  # Set model
  GibbsI = FW(y = c_expslope[,i],VAR = c_expslope$Genotype,ENV = c_expslope$Year, method="Gibbs",seed=12345,saveAt="GibbsI",nIter = 50000 ,burnIn = 5000)
  
  # Genotype Slope
  b <- data.frame(GibbsI$b)
  colnames(b) <- "b"

  # Predicted y values
  yhat <- data.frame(GibbsI$yhat)
  colnames(yhat) <- "yhat"

  # Observed y values
  y <- data.frame(GibbsI$y)
  colnames(y) <- "y"
  
  # Genotype
  g <- c_expslope$Genotype
  
  # Year
  year <- c_expslope$Year
  
  # Add column with data date to both data frames
  GDD <- colnames(c_expslope[i]) # Save as object
  b$GDD <- GDD
  
  yhat$year <- year
  yhat$Genotype <- g
  yhat$y <- y
  yhat$GDD <- GDD
  
  # Keep saving to original empty data frame
  expslope_FW_b <- rbind(expslope_FW_b, b)  
  expslope_FW_yhat <- rbind(expslope_FW_yhat, yhat)
}

# Format slope data
# Change Genotype to index to own column
expslope_FW_b  <- tibble::rownames_to_column(expslope_FW_b, "Genotype")

# Save results
write.csv(expslope_FW_b, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_AVS_b.csv", row.names = FALSE)

# Format yhat data
y_yhat <- expslope_FW_yhat

#Unnest y list from data frame
y_yhat <- unnest(y_yhat, cols = 4)

# Save results
write.csv(y_yhat, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_AVS_y.csv", row.names = FALSE)

## AVS FW Stats
AVS_FW <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/FW/results/FW_AVS_b.csv")

stats <- AVS_FW %>%
  mutate(b = b + 1) %>%
  summarise(meanFW = round(mean(b, na.rm = TRUE), 3),
            SDFW = round(sd(b, na.rm = TRUE), 3),
            CVFW = round(cv(b, na.rm = TRUE), 3),
            maxFW = round(max(b), 3),
            minFW = round(min(b, na.rm = TRUE), 3))
```





