---
title: "Static Stability: AMMI Model"
author: "Julian Cooper"
date: "11/13/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Step 0: Load packages
```{r}
library(tidyverse)
library(ggplot2)
library(agricolae)
library(tibble)
library(metan)
library(data.table)
devtools::install_github("ajaygpb/ammistability")
library(ammistability)
library(plotly)
library(hrbrthemes)
library(heatmaply)

```

# Step 1: AMMI Example and Stability for AUPC
Not included in paper
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset dataset
c_AUPC <- c[, c(1, 30:35)]

# Filter out genotypes with rep count < 5
## Need multiple reps/environments to accurately measure stability
c_AUPC <- c_AUPC %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 

############### AMMI Calculation Examples ##################

# Begin by making model using agricolae method
# Create model. Block not included, see literature (Gauch, 2012)
model <- with(c_AUPC,AMMI(Year, Genotype, Rep, AUPC, console=FALSE))

#### The following code is just a small set of examples of the possible AMMI metrics that can be produced. 

#### AMGE Sum Across Environments of GE Interaction
# AMGE: The AMGE values. The sums across environments of the GEI modeled by AMMI. The lower these parameters, the more stable a genotype is to environments. Based on sum of PC and associated eigenvalues (scalars). 
#SSI: The computed values of simultaneous selection index for yield and stability.
#rAMGE: The ranks of AMGE values.
#rY: The ranks of the mean yield of genotypes.
#means: The mean yield of the genotypes.

AMGE2 <- AMGE.AMMI(model, n=2) # Forces two PC
AMGE <- AMGE.AMMI(model) # Uses all significant PC
AVAMGE <- AVAMGE.AMMI(model) # Takes absolute value, which scales it to correlate with ASV and ASI

#### AMMI Stability Index
#ASI: The ASI values. Distance between PC1 and PC2. Higher the absolute value, the more specified a genotype is to specific environments. Conversely, lower ASI = more stable genotypes. Modified only uses significant PC.
#SSI: The computed values of simultaneous selection index for yield and stability.
#rASI: The ranks of ASI values.
#rY: The ranks of the mean yield of genotypes.
#means: The mean yield of the genotypes.

MASI <- MASI.AMMI(model) # Uses all significant PC
ASI <- ASI.AMMI(model) # Uses first two PC

#### AMMI Stability Value
#MASV: Sum of square for all significant PC. Modified AMMI Stability value. Unmodified uses first two. Lower is better.
#SSI: The computed values of simultaneous selection index for yield and stability.
#rMASV: The ranks of ASI values.
#rY: The ranks of the mean yield of genotypes.
#means: The mean yield of the genotypes.

MASV <- MASV.AMMI(model) # Uses all significant PC
ASV <- index.AMMI(model) # Uses the first two PC

# Check correlation between all possible metrics
ammistability(model, AMGE = TRUE, ASI = TRUE, ASV = TRUE, ASTAB = FALSE,
              AVAMGE = TRUE, DA = FALSE, DZ = FALSE, EV = FALSE,
              FA = FALSE, MASI = TRUE, MASV = TRUE, SIPC = FALSE,
              ZA = FALSE)
# Correlation between AVAMGE, MASV, and MASI are highest.

ammistability(model, AMGE = TRUE, ASI = TRUE, ASV = TRUE, ASTAB = TRUE,
              AVAMGE = TRUE, DA = TRUE, DZ = TRUE, EV = TRUE,
              FA = TRUE, MASI = TRUE, MASV = TRUE, SIPC = TRUE,
              ZA = TRUE)
# All correlations are good except for AMGE, but AVAMGE is in line.

############################################

# Extract results for AUPC AMMI
# Remove rep_count
c_AUPC <- c_AUPC[,1:7]

# Change to data frame
c_AUPC <- as.data.frame(c_AUPC)
str(c_AUPC)

# Make empty date frame
area_ammistability <- data.frame()

# For each column with area data
for (i in 7:7){
  
  # set model
  model <- with(c_AUPC, AMMI(Year, Genotype, Rep, c_AUPC[,i], console=FALSE))
  
  # make new data frame
  area_rubberduck <- data.frame()
  
  # Extract GDD
  GDD <- colnames(c_AUPC[i])
   
  #AMMI metrics give the give the following error:
  # Error in gamma.n %*% diag(lambda.n) %*% t(delta.n) : 
  # non-conformable arguments
  ### Happens when there is only one significant PC. Manually set n=2 if code occurs.
  
  # Calculate AVAMGE
  AVAMGE <- try(AVAMGE.AMMI(model))
  if (class(AVAMGE) == "try-error") { # if error
    AVAMGE <- AVAMGE.AMMI(model, n=2) # use 2 PC
  }
  colnames(AVAMGE) <- c("AVAMGE", "AVAMGE_SSI", "rAVAMGE", "rY", "means") # change columns names
  AVAMGE  <- tibble::rownames_to_column(AVAMGE, "Genotype") # make row names a new column

  # Calculate MASI
  MASI <- try(MASI.AMMI(model))
  if (class(MASI) ==  "try-error") { # if error
    MASI <- MASI.AMMI(model, n=2) # use 2 PC
  }
  colnames(MASI) <- c("MASI", "MASI_SSI", "rMASI", "rY", "means") # change columns names
  MASI  <- tibble::rownames_to_column(MASI, "Genotype") # make row names a new column
  
  # Calculate MASV
  MASV <- try(MASV.AMMI(model))
  if (class(MASV) =="try-error") { # if error
    MASV <- MASV.AMMI(model, n=2) # use 2 PC
  }
  colnames(MASV) <- c("MASV", "MASV_SSI", "rMASV", "rY", "means") # change columns names
  MASV  <- tibble::rownames_to_column(MASV, "Genotype") # make row names a new column
  
  # Merge results into new data frame
  area_rubberduck <- merge(AVAMGE, MASI, by = "Genotype")
  area_rubberduck <- merge(area_rubberduck, MASV, by = "Genotype")
  area_rubberduck <- cbind(area_rubberduck, GDD)

  area_ammistability <- rbind(area_ammistability, area_rubberduck)

}

# Clean data frame by removing duplicated columns
area_ammistability_clean <- area_ammistability[, c(1:9, 12:14, 17)]

# Save data frame
write.csv(area_ammistability_clean, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_AUPC.csv", row.names = FALSE)

## AUPC AVAMGE Stats
AUPC_ammi <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_AUPC.csv")

stats <- AUPC_ammi %>%
  summarise(mean = round(mean(AVAMGE, na.rm = TRUE), 3),
            sd = round(sd(AVAMGE, na.rm = TRUE), 3),
            CV = round(cv(AVAMGE, na.rm = TRUE), 3),
            max = round(max(AVAMGE, na.rm = TRUE), 3),
            min = round(min(AVAMGE, na.rm = TRUE),3))
```

# Step 2: Ammistability for Point
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset data
c_point <- c[, 1:34]

# Make empty date frame
point_ammistability <- data.frame()

# For each column with point data
for (i in 2:29){ 
  
  # Make new data farme
  c_model <- c_point[,c(1, i, 30:34)]
  
  # Remove missing
  c_model <- na.omit(c_model)
  
  # Filter genotypes with less than 5 reps
  c_model <- c_model %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 
  
  # Make sure data is in data frame form
  c_model <- as.data.frame(c_model)
  #str(c_model)
  
  # Set model
  model2 <- with(c_model,AMMI(Year, Genotype, Rep, c_model[,2], console=FALSE))
  
  point_rubberduck <- data.frame() # make new empty data frame
  
  GDD <- colnames(c_model[2]) # Extract GDD
   
  #AMMI metrics give the give the following error:
  # Error in gamma.n %*% diag(lambda.n) %*% t(delta.n) : 
  # non-conformable arguments
  ### Happens when there is only one significant PC. Manually set n=2 if code occurs.
  
  # Calculate AVAMGE, MASI, and MASV, using 2 PC if error
  AVAMGE <- try(AVAMGE.AMMI(model2))
  if (class(AVAMGE) == "try-error") {
    AVAMGE <- AVAMGE.AMMI(model2, n=2)
  }
  colnames(AVAMGE) <- c("AVAMGE", "AVAMGE_SSI", "rAVAMGE", "rY", "means")
  AVAMGE  <- tibble::rownames_to_column(AVAMGE, "Genotype")

  MASI <- try(MASI.AMMI(model2))
  if (class(MASI) ==  "try-error") {
    MASI <- MASI.AMMI(model2, n=2)
  }
  colnames(MASI) <- c("MASI", "MASI_SSI", "rMASI", "rY", "means")
  MASI  <- tibble::rownames_to_column(MASI, "Genotype")
  
  MASV <- try(MASV.AMMI(model2))
  if (class(MASV) =="try-error") {
    MASV <- MASV.AMMI(model2, n=2)
  }
  
  colnames(MASV) <- c("MASV", "MASV_SSI", "rMASV", "rY", "means") # Set column names
  MASV  <- tibble::rownames_to_column(MASV, "Genotype") # Change row names to column
  
  # Merge all and make new data frame
  point_rubberduck <- merge(AVAMGE, MASI, by = "Genotype")
  point_rubberduck <- merge(point_rubberduck, MASV, by = "Genotype")
  point_rubberduck <- cbind(point_rubberduck, GDD)

  point_ammistability <- rbind(point_ammistability, point_rubberduck)

}

# Clean data
point_ammistability_clean <- point_ammistability[, c(1:9, 12:14, 17)]

# Save data frame
write.csv(point_ammistability_clean, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_point.csv", row.names = FALSE)

############ Stats for % CC ##################
point_AMMI <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_point.csv")
point_AMMI <- point_AMMI[,c(2, 13)]

## Average AVAMGE at each GDD
supf4 <- point_AMMI %>%
  group_by(GDD) %>%
  summarise(meanAVAMGE = round(mean(AVAMGE, na.rm = TRUE), 3),
            SDAVAMGE = round(sd(AVAMGE, na.rm = TRUE), 3),
            CV = round(cv(AVAMGE, na.rm = TRUE), 3),
            minAVAMGE = round(min(AVAMGE, na.rm = TRUE), 3),
            maxAVAMGE = round(max(AVAMGE, na.rm = TRUE), 3))

## Average AVAMGE at each Growth Stage
point_AMMI %>%
  mutate(Stage = case_when(GDD %in% c(seq(300,450)) ~ "Juv.",
                           GDD %in% c(seq(500,1100)) ~ "Adult",
                           GDD %in% c(seq(1150,1650)) ~"Rep.")) %>%
  group_by(Stage) %>%
  summarise(meanAVAMGE = round(mean(AVAMGE, na.rm = TRUE), 3),
            SDAVAMGE = round(sd(AVAMGE, na.rm = TRUE), 3),
            CV = round(cv(AVAMGE, na.rm = TRUE), 3),
            minAVAMGE = round(min(AVAMGE, na.rm = TRUE), 3),
            maxAVAMGE = round(max(AVAMGE, na.rm = TRUE), 3))

```

# Step 3: Ammistability for Slope
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset data
c_slope <- c[, c(1, 30:34, 36:62)]

# Make empty date frame
slope_ammistability <- data.frame()

# For each column with slope data
for (i in 7:33){
  
  c_model <- c_slope[,c(1:6, i)]
  
  c_model <- na.omit(c_model)
  
  # Filter genotypes with less than 5 reps
  c_model <- c_model %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 
  
  # Make sure data is in data frame form
  c_model <- as.data.frame(c_model)
  #str(c_model)
  
  # Set model
  model3 <- with(c_model,AMMI(Year, Genotype, Rep, c_model[,7], console=FALSE))
  
  slope_rubberduck <- data.frame() # make empty data frame
  
  GDD <- colnames(c_model[7]) # extract GDD
  
  #AMMI metrics give the give the following error:
  # Error in gamma.n %*% diag(lambda.n) %*% t(delta.n) : 
  # non-conformable arguments
  ### Happens when there is only one significant PC. Manually set n=2 if code occurs.
  
  # Calculate AVAMGE, MASI, and MASV, using 2 PC if error
  
  AVAMGE <- try(AVAMGE.AMMI(model3))
  if (class(AVAMGE) == "try-error") {
    AVAMGE <- AVAMGE.AMMI(model3, n=2)
  }
  colnames(AVAMGE) <- c("AVAMGE", "AVAMGE_SSI", "rAVAMGE", "rY", "means")
  AVAMGE  <- tibble::rownames_to_column(AVAMGE, "Genotype")

  MASI <- try(MASI.AMMI(model3))
  if (class(MASI) ==  "try-error") {
    MASI <- MASI.AMMI(model3, n=2)
  }
  colnames(MASI) <- c("MASI", "MASI_SSI", "rMASI", "rY", "means")
  MASI  <- tibble::rownames_to_column(MASI, "Genotype")
  
  MASV <- try(MASV.AMMI(model3))
  if (class(MASV) =="try-error") {
    MASV <- MASV.AMMI(model3, n=2)
  }
  colnames(MASV) <- c("MASV", "MASV_SSI", "rMASV", "rY", "means")
  MASV  <- tibble::rownames_to_column(MASV, "Genotype")

  # Merge all and make new data frame
  slope_rubberduck <- merge(AVAMGE, MASI, by = "Genotype")
  slope_rubberduck <- merge(slope_rubberduck, MASV, by = "Genotype")
  slope_rubberduck <- cbind(slope_rubberduck, GDD)  
  
  slope_ammistability <- rbind(slope_ammistability, slope_rubberduck)

}

# Clean data
slope_ammistability_clean <- slope_ammistability[, c(c(1:9, 12:14, 17))]

# Save data frame
write.csv(slope_ammistability_clean, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_slope.csv", row.names = FALSE)

## AVS AVAMGE Stats
slope_ammi <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_slope.csv")

# For each GDD
stats <- slope_ammi %>%
  group_by(GDD) %>%
  summarise(mean = round(mean(AVAMGE, na.rm = TRUE), 3),
            sd = round(sd(AVAMGE, na.rm = TRUE), 3),
            CV = round(cv(AVAMGE, na.rm = TRUE), 3),
            max = round(max(AVAMGE, na.rm = TRUE), 3),
            min = round(min(AVAMGE, na.rm = TRUE),3))

#For each growth stage
# Add Phase
juvenile <- c("300-350", "350-400", "400-450")
adult <- c("450-500", "500-550", "550-600", "600-650", "650-700", "700-750",  "750-800",  "800-850",  "850-900",  "900-950",  "950-1000",  "1000-1050", "1050-1100",  "1100-1150")
reproductive <- c("1150-1200",  "1200-1250",  "1250-1300",  "1300-1350",  "1350-1400",  "1400-1450",  "1450-1500",  "1500-1550",  "1550-1600",  "1600-1650")

# Average, Max, Min, and SD Slope of each Growth Stage
stats <- slope_ammi %>%
  mutate(
    GDD = case_when(
      GDD == "1" ~ "300-350",
      GDD == "2" ~ "350-400",
      GDD == "3" ~ "400-450",
      GDD == "4" ~ "450-500",
      GDD == "5" ~ "500-550",
      GDD == "6" ~ "550-600",
      GDD == "7" ~ "600-650",
      GDD == "8" ~ "650-700",
      GDD == "9" ~ "700-750",
      GDD == "10" ~ "750-800",
      GDD == "11" ~ "800-850",
      GDD == "12" ~ "850-900",
      GDD == "13" ~ "900-950",
      GDD == "14" ~ "950-1000",
      GDD == "15" ~ "1000-1050",
      GDD == "16" ~ "1050-1100",
      GDD == "17" ~ "1100-1150",
      GDD == "18" ~ "1150-1200",
      GDD == "19" ~ "1200-1250",
      GDD == "20" ~ "1250-1300",
      GDD == "21" ~ "1300-1350",
      GDD == "22" ~ "1350-1400",
      GDD == "23" ~ "1400-1450",
      GDD == "24" ~ "1450-1500",
      GDD == "25" ~ "1500-1550",
      GDD == "26" ~ "1550-1600",
      GDD == "27" ~ "1600-1650")
  ) %>% 
  mutate(Stage = case_when(GDD %in% juvenile ~ "Juvenile",
                           GDD %in% adult ~ "Adult",
                           GDD %in% reproductive ~"Reproductive")) %>%
  group_by(Stage) %>%
 summarise(mean = round(mean(AVAMGE, na.rm = TRUE), 3),
            sd = round(sd(AVAMGE, na.rm = TRUE), 3),
            CV = round(cv(AVAMGE, na.rm = TRUE), 3),
            max = round(max(AVAMGE, na.rm = TRUE), 3),
            min = round(min(AVAMGE, na.rm = TRUE),3))

```

# Step 4: Ammistability for AVS
Not included in paper
```{r eval=FALSE}

# Load data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Subset data
c_AVS <- c[, c(1, 30:34, 63)]

# Make empty date frame
AVS_ammistability <- data.frame()

# For each column with slope data
for (i in 7:7){
  
  c_model <- c_AVS[,c(1:6, i)]
  
  c_model <- na.omit(c_model)
  
  # Filter genotypes with less than 5 reps
  c_model <- c_model %>%
  group_by(Genotype) %>%
  mutate(rep_count = n()) %>%
  ungroup() %>%
  filter(rep_count > 4) 
  
  # Make sure data is in data frame form
  c_model <- as.data.frame(c_model)
  #str(c_model)
  
  # Set model
  model3 <- with(c_model,AMMI(Year, Genotype, Rep, c_model[,7], console=FALSE))
  
  slope_rubberduck <- data.frame() # new empty data frame
  
  GDD <- colnames(c_model[7]) # extract GDD
  
   #AMMI metrics give the give the following error:
  # Error in gamma.n %*% diag(lambda.n) %*% t(delta.n) : 
  # non-conformable arguments
  ### Happens when there is only one significant PC. Manually set n=2 if code occurs.
  
  # Calculate AVAMGE, MASI, and MASV, using 2 PC if error
  
  AVAMGE <- try(AVAMGE.AMMI(model3))
  if (class(AVAMGE) == "try-error") {
    AVAMGE <- AVAMGE.AMMI(model3, n=2)
  }
  colnames(AVAMGE) <- c("AVAMGE", "AVAMGE_SSI", "rAVAMGE", "rY", "means")
  AVAMGE  <- tibble::rownames_to_column(AVAMGE, "Genotype")

  MASI <- try(MASI.AMMI(model3))
  if (class(MASI) ==  "try-error") {
    MASI <- MASI.AMMI(model3, n=2)
  }
  colnames(MASI) <- c("MASI", "MASI_SSI", "rMASI", "rY", "means")
  MASI  <- tibble::rownames_to_column(MASI, "Genotype")
  
  MASV <- try(MASV.AMMI(model3))
  if (class(MASV) =="try-error") {
    MASV <- MASV.AMMI(model3, n=2)
  }
  colnames(MASV) <- c("MASV", "MASV_SSI", "rMASV", "rY", "means")
  MASV  <- tibble::rownames_to_column(MASV, "Genotype")
  
  # Merge all and make new data frame

  slope_rubberduck <- merge(AVAMGE, MASI, by = "Genotype")
  slope_rubberduck <- merge(slope_rubberduck, MASV, by = "Genotype")
  slope_rubberduck <- cbind(slope_rubberduck, GDD)  
  
  AVS_ammistability <- rbind(AVS_ammistability, slope_rubberduck)

}

# Clean data
AVS_ammistability_clean <- AVS_ammistability[, c(c(1:9, 12:14, 17))]

# Save data frame
write.csv(AVS_ammistability_clean, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_AVS.csv", row.names = FALSE)

## AVS AVAMGE Stats
AVS_ammi <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_AVS.csv")

stats <- AVS_ammi %>%
  summarise(mean = round(mean(AVAMGE, na.rm = TRUE), 3),
            sd = round(sd(AVAMGE, na.rm = TRUE), 3),
            CV = round(cv(AVAMGE, na.rm = TRUE), 3),
            max = round(max(AVAMGE, na.rm = TRUE), 3),
            min = round(min(AVAMGE, na.rm = TRUE),3))
```

# Step 5: Rank stability of genotypes and compare across traits
## Step 5A: Merge data
```{r}

# Load and format all AMMI data frames
point <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_point.csv")
point_formatted <- point[,c(1,4,13)]
point_formatted$Trait <- "Point"

slope <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_slope.csv")
slope_formatted <- slope[,c(1,4,13)]
slope_formatted$Trait <- "Slope"

AVS <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_AVS.csv")
AVS_formatted <- AVS[,c(1,4,13)]
AVS_formatted$Trait <- "AVS"

AUPC <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_AUPC.csv")
AUPC_formatted <- AUPC[,c(1,4,13)]
AUPC_formatted$Trait <- "AUPC"

# Merge df
AMMI_all <- rbind(point_formatted, slope_formatted, AVS_formatted, AUPC_formatted)

# Add trait-GDD ID
AMMI_all <- AMMI_all %>%
  mutate(ID = paste(GDD, Trait, sep = "_")) %>%
  reframe(Genotype, rAVAMGE, ID)

# Save data frame
write.csv(AMMI_all, "/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_all.csv", row.names = FALSE)

```

## Step 5B: Genotype stability rank correlations between traits
```{r}

# Load data
AMMI_all <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_all.csv")

# Pivot wide
AMMI_wide <- AMMI_all %>%
  pivot_wider(names_from = ID, values_from = rAVAMGE)

# Correlations between geno ranks
AMMI_wide %>%
  summarise(cor(AUPC_AUPC, `1450_Point`, use = "complete.obs"))


```

## Step 5C: How does stability correlate with performance (terminal CC %)?
```{r}

# Load % CC stability data
a <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/4_trait_stability/AMMI/results/AMMI_point.csv")

# Get average AMMI for each geno
## LOW = MORE STABLE. According to Bernardo 2020, more stable should have worse overall trait performance
ammi_point <- a %>%
  group_by(Genotype) %>%
  summarise(avg_AMMI = mean(AVAMGE))

# Load trait data
c <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/combined_years/results/canopy_cover_AUPC_slope_AVS.csv", check.names = FALSE)

# Trim % CC only
c_point <- c[c(2:29, 32)]

c_point_avg <- c_point %>%
  pivot_longer(cols = c(1:28), names_to = "GDD", values_to = "CC") %>%
  group_by(Genotype, GDD) %>%
  summarise(avg_CC = mean(CC, na.rm = TRUE)) %>%
  pivot_wider(names_from = "GDD", values_from = "avg_CC")

# Merge
ammi_c <- merge(ammi_point, c_point_avg, by = "Genotype")

# How does avg stability correlate with terminal CC?
ammi_c %>%
  summarise(cor(avg_AMMI, `1450`))

# How does terminal stability correlate with terminal CC?
a_wide <- a %>%
  reframe(Genotype, AVAMGE, GDD) %>%
  pivot_wider(names_from = GDD, values_from = AVAMGE)

cor(a_wide$`1450`, ammi_c$`1450`, use = "complete.obs")



```