---
title: "2019 Canopy Cover"
author: "Julian Cooper"
date: "11/7/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Step 0: Load libraries
```{r load_libraries}

library(dplyr)
library(data.table)  
library(tidyverse)

```

# Step 1: Combine canopy cover trait files
```{r combine_files, results='hide'}

# Get list of files to combine
txt_files_ls <- list.files("/Users/jcooper/Desktop/thesis_research/canopy_cover/1_image_masking/image_analysis/data_analysis/2019_doubledilated/canopy_closure/",pattern = "txt",full.names = T, recursive = TRUE)

# Turn into data frame
txt_files_df <- lapply(txt_files_ls, function(x) {read.table(file = x, header = F, col.names = x, sep =",")})

# Combine files
combined_df <- do.call("cbind", lapply(txt_files_df, as.data.frame)) 

## Rename columns
# Make list of file column names
old_names <- c(colnames(combined_df))

# Make list of new GDD column names
# First flight is technically 302 GDD. Round down to 300 so loess can span 300-1450 with 2020 and 2021.
new_names <- c(300, 428.5, 491, 583, 722, 769.5, 811, 1115, 1230.5, 1418, 1650.5)

# Rename columns to GDD 
for(i in 1:11) names(combined_df)[names(combined_df) == old_names[i]] = new_names[i]

# Save as .csv
write.csv(combined_df,"/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/1PlantRatio.csv", row.names = FALSE)

```

# Step 2: Generate plot order from QGIS ROI files
```{r plot_order }
################ PRODUCES PLOT OUTPUT ORDER BY DETERMINING PLOTS NORTH TO SOUTH THEN WEST TO EAST ######################
######## CHANGE INPUTS ########################

# read in plot boundaries and coordinates
plot_roi <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/plot_order/Plots_widiv_ROI2019.csv")

# read in list of plots
plot_list <- as.matrix(read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/plot_order/Plots_only_WIDIV.csv"))

# Path of location to save the plot orders
plot_order_loc = "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/plot_order/"

# Name of Field
Field = 'WIDIV'
# give the dimensions of the plot grid
plot_grid_width =  48    # number of plots in a range
plot_grid_length =  22   # number of ranges
total_num_plots = 1056

################################################################################################################################
options(digits = 15) # change the number of digits seen in files
# remove unnecessary columns
plot_roi$Geometry <- NULL
plot_roi$BoundingBox <- NULL
plot_roi$left <- NULL
plot_roi$top <- NULL
plot_roi$right <- NULL
plot_roi$bottom <- NULL

# split x column 
plot_roi <- splitstackshape::cSplit(plot_roi, "X", sep = ",")

# remove extra X columns
plot_roi$X_6 <- NULL
plot_roi$X_5 <- NULL
plot_roi$X_4 <- NULL
plot_roi$X_3 <- NULL
plot_roi$X_2 <- NULL

# remove the bracket from X_1
plot_roi <- splitstackshape::cSplit(plot_roi, "X_1", sep = "[")

# rename keep row and remove unnecessary rows
plot_roi$X_coor <- plot_roi$X_1_2
plot_roi$X_1_1 <- NULL
plot_roi$X_1_2 <- NULL

# Split y column, remove extra Y columns, remove the bracket, and rename the keep row
plot_roi <- splitstackshape::cSplit(plot_roi, "Y", sep = ",")
plot_roi <- splitstackshape::cSplit(plot_roi, "Y_1", sep = "[")
plot_roi$Y_coor <- plot_roi$Y_1_2
plot_roi$Y_1_1 <- NULL
plot_roi$Y_1_2 <- NULL
plot_roi$Y_6 <- NULL
plot_roi$Y_5 <- NULL
plot_roi$Y_4 <- NULL
plot_roi$Y_3 <- NULL
plot_roi$Y_2 <- NULL

# Organize the plots X increasing and Y decreasing
plot_roi_temp <- plot_roi[order(plot_roi$X_coor),]
plot_roi.2 <- data.frame()
for(i in 1:plot_grid_length){
  if(nrow(plot_roi_temp)>=plot_grid_width){
    plot_roi_temp2 <- plot_roi_temp[1:plot_grid_width,]
  } else{
    plot_roi_temp2 <- plot_roi_temp[1:nrow(plot_roi_temp),]
  }
  plot_roi_temp2 <- plot_roi_temp2[order(plot_roi_temp2$Y_coor, decreasing = T),]
  plot_roi_temp <- plot_roi_temp[(plot_grid_width +1):nrow(plot_roi_temp),]
  
  plot_roi.2 <- rbind(plot_roi.2, plot_roi_temp2)
}

# add new column with non-serpentine order
plot_roi.2$non_serp <- 1:(total_num_plots) 

# reorder the list of plots so no longer serpentine
Count = 1  # setup for 'for loop'
plot_list2 = matrix(nrow = (total_num_plots), ncol = 1) # make a new plot list file to put the new order in

for (Count in 1:plot_grid_length){
  if (Count %% 2 == 1){
    for (plot in ((((Count - 1)* plot_grid_width)+1):(Count *plot_grid_width))){
      plot_list2[plot,1] = plot_list[plot,1]}
    Count = Count + 1
  }else if (Count %% 2 == 0){
    for (plot in ((((Count - 1)* plot_grid_width)+1):(Count *plot_grid_width))){
      if ((plot %% plot_grid_width) != 0){
        plot_list2[((Count * plot_grid_width) - ((plot %% plot_grid_width)-1)),1] = plot_list[plot,1]
      }else if ((plot %% plot_grid_width) == 0){ 
        plot_list2[(plot - (plot_grid_width -1)),1] = plot_list[plot,]}}
    Count = Count + 1
  }
}

# add non-serpentine plot order column to new plot_list2 so will have common column with plot_roi.2
plot_list2 <- as.data.frame(plot_list2)
plot_list2$non_serp <- 1:(total_num_plots)

# add new plot order list to plot_roi.2
plot_roi.3 <- merge(plot_roi.2, plot_list2, by = "non_serp")

## reorganize back into original roi order (as output would be -> ordered by fields)
plot_order.4 <- plot_roi.3[order(Fields)]
plot_order.4$Plot <- plot_order.4$V1

plot_order.5 <- plot_order.4$Plot

#Print out the Plot Order text file
write.table(plot_order.4$Plot, file = paste0(plot_order_loc,"Plots_",Field,"_Ordered.txt"), sep = "\n", row.names = F, col.names = "Plot")

```

# Step 3: Add plot order to trait data frame
```{r experimental_design_merge}

# Read in final Plant Ratio file
CC_final <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/1PlantRatio.csv", check.names = FALSE)

# Read in plot order file
p_order <- read.table("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/plot_order/Plots_WIDIV_Ordered.txt", header = TRUE)

#Column bind data frames
CC_final1 <- cbind(p_order, CC_final) 

#Filter out PH207 filler rows
CC_final2 <- CC_final1[CC_final1$Plot !='PH207',]

write.csv(CC_final2, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/2PlantRatio_plot.csv",  row.names = FALSE)

##################
# QC

# Pivot Table into "long"
ccfinal_long <- pivot_longer(CC_final2, cols=2:12, names_to="flight", values_to="canopy_cover")

# Make "flight" numeric
ccfinal_long$flight <- as.numeric(ccfinal_long$flight)

# Plot
ggplot(data=ccfinal_long, aes(x=flight, y=canopy_cover)) +
  #geom_point(color="gray", size=.1) +
  geom_line(aes(group=Plot), color="gray") +
  stat_smooth(method="loess", formula = y ~ x) +
  stat_summary(aes(flight, canopy_cover), geom = 'point', fun = mean, shape = 17, size = 3, col = "red") + 
  theme_bw() + 
  xlab('Flight (GDD)') + 
  ylab('Plant Ratio') + 
  ggtitle("2019 Double Dilated Percent Canopy Cover")+
  labs(caption = "n = 1040")


```

# Step 4: Adjust for stand
```{r adjust_stand}

# Read in PlantRatio_plot
c2019 <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/2PlantRatio_plot.csv", check.names = FALSE)

# Read in stand counts
stand2019 <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/field_data/2019_stand_JC01052022.csv")

# Merge by "Plot"
c_stand <- merge(c2019, stand2019, by = "Plot")

# Change stand counts to numeric
c_stand$Stand <- as.numeric(c_stand$Stand)

# Filter for stand counts > 14
c_standfilter <- c_stand %>%
  filter(Stand > 14) # n = 944

# Save data frame
write.csv(c_standfilter, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/3PlantRatio_plot_stand.csv",  row.names = FALSE)

##################
# QC

# Pivot Table into "long"
ccfinal_long <- pivot_longer(c_standfilter, cols=2:12, names_to="flight", values_to="canopy_cover")

# Make "flight" numeric
ccfinal_long$flight <- as.numeric(ccfinal_long$flight)

# Plot
ggplot(data=ccfinal_long, aes(x=flight, y=canopy_cover)) +
  #geom_point(color="gray", size=.1) +
  geom_line(aes(group=Plot), color="gray") +
  stat_smooth(method="loess", formula = y ~ x) +
  stat_summary(aes(flight, canopy_cover), geom = 'point', fun = mean, shape = 17, size = 3, col = "red") + 
  theme_bw() + 
  xlab('Flight (GDD)') + 
  ylab('Plant Ratio') + 
  ggtitle("2019 Double Dilated Percent Canopy Cover - Weeds, Stand") +
  labs(caption = "n = 944")

```

# Step 5: Filter peaks and valleys
Check /Users/jcooper/Desktop/thesis_research/canopy_cover_master/archive/loess/2019_doubledilated/2019_dipfilterfun_12102021.Rmd. Remove flight Original 766 / New GDD 769.5 to minimize data loss. 
```{r filter_pv}

# Read in data
c2019 <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/3PlantRatio_plot_stand.csv", check.names = FALSE)

# Remove extra columns
c2019 <- c2019[, 1:12]

# Remove bad flight date 769.5 GDD
c2019 <- c2019[,-7]

# Function to filter peaks and valleys
big_data <- data.frame() #make empty data frame
filter_peaks <- function(df) { #Input df into function
  for (i in seq_len(nrow(df))) { #For each row in data frame
  row <- df[i,] #Subset row
  names <- colnames(row) # Make list of column names
  # Pivot row into long
  row <- pivot_longer(row, cols = names[2:11], names_to = "flight", values_to = "closure") 
  # print(row)
  
  left <- 1 # Set left end of sliding evaluation window
  right <- 3 # Set right end of sliding evaluation window
  
  # Check for dips 
  for (j in 1:1) {row[j,4] <- NA} # Nothing to compare first flight to, automatically label good
  
  for (j in 2:9) { #For each column in row starting at 3
    
    # Rules of finding valleys
    if (row[j, 3] < row[left, 3] * .90) 
      if (row[right, 3] > row[j, 3])
        if (row[left, 3] < row[right, 3])
        {row[j,4] <- "dip"}
    
    # Rules for finding peaks
    if (row[j, 3] > row[right, 3] * 1.10) 
      if (row[right, 3] > row[left, 3])
        if (row[left, 3] < row[j, 3])
        {row[j,4] <- "peak"}
    
    # Move sliding window to next data point
    left <- left + 1
    right <- right + 1
  }
  
  # Rules for checking if last data point is valley, can't check for peaks b/c nothing to compare to on right
  for (j in 10:10) {
    if ((row[j, 3]/row[9, 3]) < .90) {row[j,4] <- "dip"} else {row[j,4] <- NA}

  }
  big_data <- rbind(big_data, row) # Bind each row back to the empty data frame created outside the function

  }
  c2019_dipdrop <<- big_data # Export data frame with peaks and valleys labeled outside of function

}

# Run Function  
filter_peaks(c2019)

# Make new column copying closure
c2019_dipdrop$closureNA <- c2019_dipdrop$closure

# Replace data points with NA if tagged as peak or valley
c2019_dipdrop$closureNA <- ifelse(is.na(c2019_dipdrop$...4), c2019_dipdrop$closure, ifelse(c2019_dipdrop$...4 == "dip",  NA, ifelse(c2019_dipdrop$...4 == "peak", NA, c2019_dipdrop$closure)))

# Remove extra columns
c2019_dipdrop <- c2019_dipdrop[,c(1, 2, 5)]

# Pivot wide To count # of NA per row
c2019_wide <- pivot_wider(c2019_dipdrop, names_from = "flight", values_from = "closureNA")

# Make new column with na counts
c2019_wide$na_count <- apply(c2019_wide, 1, function(x) sum(is.na(x)))

# Pivot back to long to filter out na counts past certain threshold
c2019_long <- c2019_wide %>%
  pivot_longer(cols = 2:11, names_to = "flight", values_to = "closure")

# Count na
c2019_long %>%
  summarise(sum(is.na(closure))) # 721 NA data points

# Filter plots with na count > 3
c2019_filter <- c2019_long %>%
  filter(na_count < 3)

# Count number of plots
c2019_filter %>%
  summarise(n_distinct(Plot)) # Keeps 855 of 941 plots

# Count na
c2019_filter%>%
  summarise(sum(is.na(closure))) # 404 NA data points

# Pivot back to wide
c2019_final <- pivot_wider(c2019_filter, names_from = "flight", values_from = "closure")

# Save data frame
write.csv(c2019_final, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/4PlantRatio_plot_stand_dips.csv",  row.names = FALSE)

##################
# QC

# Make "flight" numeric
c2019_filter$flight <- as.numeric(c2019_filter$flight)

# Plot
ggplot(data=c2019_filter, aes(x=flight, y=closure)) +
  #geom_point(color="gray", size=.1) +
  geom_line(data=c2019_filter[!is.na(c2019_filter$closure),], aes(group=Plot), color="gray") +
  stat_smooth(method="loess", formula = y ~ x) +
  stat_summary(aes(flight, closure), geom = 'point', fun = mean, shape = 17, size = 3, col = "red") + 
  theme_bw() + 
  xlab('Flight (GDD)') + 
  ylab('Plant Ratio') + 
  ggtitle("2019 Double Dilated Percent Canopy Cover - Weeds, Stand, Peaks and Valleys") +
  labs(caption = "n = 855")


```

# Step 6: Determine best span
```{r span, eval = FALSE}

C <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/4PlantRatio_plot_stand_dips.csv", check.names = FALSE)

# Drop na_count column
C <- C[, c(1, 3:12)]

column_names <- colnames(C)

C_long <- pivot_longer(C, column_names[2:11], names_to = "flight", values_to = "canopy_cover") #Pivot long based on flight columns

C_long$flight <- as.numeric(C_long$flight)

#Determine best span
span.seq <- seq(from = 0.15, to = 0.95, by = 0.05) #explores range of spans
k <- 100 #number of folds
set.seed(1) # replicate results
folds <- sample(x = 1:k, size = nrow(C_long), replace = TRUE)
cv.error.mtrx <- matrix(rep(x = NA, times = k * length(span.seq)), 
                        nrow = length(span.seq), ncol = k)

for(i in 1:length(span.seq)) {
  for(j in 1:k) {
    loess.fit <- loess(formula = canopy_cover ~ flight, data = C_long[folds != j, ], span = span.seq[i])
    preds <- predict(object = loess.fit, newdata = C_long[folds == j, ])
    cv.error.mtrx[i, j] <- mean((C_long$canopy_cover[folds == j] - preds)^2, na.rm = TRUE)
    # some predictions result in `NA` because of the `x` ranges in each fold
 }
}

# mean error from each fold
cv.errors <- rowMeans(cv.error.mtrx)

# determine which span was the best span
best.span.i <- which.min(cv.errors)
best.span.i
best_span <- span.seq[best.span.i]

plot(x = span.seq, y = cv.errors, type = "l", main = "CV Plot")
points(x = span.seq, y = cv.errors, pch = 20, cex = 0.75, col = "blue")
points(x = span.seq[best.span.i], y = cv.errors[best.span.i], pch = 20, cex = 1, col = "red")

# Best span = .15 - .35. Set eval = FALSE when knitting because code takes so long to run.
```

# Step 7: Generate Loess curves
```{r loess, results = "hide", warning = FALSE}

# Read in data
C <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/4PlantRatio_plot_stand_dips.csv", check.names = FALSE)

# Set span to lowest number that will still run. .25.
# Set span
best_span <- .25

# Drop na_count column
C <- C[, c(1, 3:12)]

# Loess loop
for(i in 1:nrow(C)){ # For i in 1:number of rows in df C
  
  #Format Data
  plot <- C[i,] #Subset row i
  plot_long <- pivot_longer(plot, cols=2:11, names_to="flight", values_to="canopy_cover") #Pivot row i to long
  plot_long$flight <- as.numeric(plot_long$flight) #Make flight numeric
  plot_name <- as.matrix(C[i,1]) #Extract plot name
  
  #Run Model
  plot.loess <- loess(canopy_cover ~ flight, plot_long, span = best_span) #Run the model
  
  #Predict Along Curve
  predictions <- data.frame(seq(300,1650, by = 50)) #Make data frame with column of GDD at intervals 1-1650, by 50
  predictions$canopy_cover <- predict(plot.loess, data.frame(flight = seq(300, 1650, by = 50))) #Predict for each point along observed GDD curve
  
  #Format Predictions
  predictions$GDD <- predictions[,1] #Rename column of 1-1650 sequence as GDD
  predictions[,1] <- NULL #Remove unnamed column of sequence
  predictions <- predictions[, c(2,1)] #Reorganize so GDD is first column, canopy_cover is second
  # #Plot Predictions
  # ggplot(predictions, aes(x = GDD, y = canopy_cover)) +
  #   geom_point() +
  #   geom_line()
  
  # Rename canopy_cover column to the name of the plot
  colnames(predictions)[which(colnames(predictions)=="canopy_cover")] <- plot_name
  
  # Make a file with Plot predictions in order to keep row names and not export the GDD repeatedly
  Plot_Predictions <- predictions
  Plot_Predictions$GDD <- NULL

  Plot_Predictions <- (t(Plot_Predictions)) #transpose
  
  # Export column names for first row only
  if(i==1){
    y <- Plot_Predictions 
    
  } else {
    y <- rbind(y, Plot_Predictions) #Doesn't export name after i ==1

  }
}

#Format Predictions
colnames(y) <- seq(300, 1650, by = 50) # Add column names
y2 <- as.data.frame(y, row.names = NULL) # Change to data frame
y3 <- tibble::rownames_to_column(y2, ) # Make row names (Plot) a column

Loess_Plot_Predictions2019 <- y3 # Rename data frame
Loess_Plot_Predictions2019$Plot <- Loess_Plot_Predictions2019$rowname # Copt row name column with new name
Loess_Plot_Predictions2019$rowname <- NULL # Remove old row name column

Loess_Plot_Predictions2019 <- Loess_Plot_Predictions2019[, c(29,1:28)] #Reorganize so GDD is first column, canopy_cover is second

# Save data frame
write.csv(Loess_Plot_Predictions2019, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/5PlantRatio_loess.csv",  row.names = FALSE)

```

# Step 8: Merge Loess curves with experimental design
```{r design}

# Load loess
loess <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/data_processing/5PlantRatio_loess.csv", check.names = FALSE)

# Load experimental design info
design <- read.csv("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/data/field_data/2019_Widiv_FieldBook.csv")

# Merge data frames
design_loess <- merge(loess, design, by = "Plot")

write.csv(design_loess, "/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/results/2019_loess_CC.csv",  row.names = FALSE)

##################
# QC

# Pivot Table into "long"
c_long <- pivot_longer(design_loess, cols=2:29, names_to="flight", values_to="closure")

# Make "flight" numeric
c_long$flight <- as.numeric(c_long$flight)

# Plot
p <- ggplot(data=c_long, aes(x=flight, y=closure)) +
  #geom_point(color="gray", size=.1) +
  geom_line(aes(group=Plot), color="gray") +
  stat_smooth(method="loess", formula = y ~ x) +
  stat_summary(aes(flight, closure), geom = 'point', fun = mean, shape = 17, size = 3, col = "red") + 
  theme_bw() + 
  xlab('Flight (GDD)') + 
  ylab('Plant Ratio') + 
  ggtitle("2019 Double Dilated Percent Canopy Cover - Loess") +
  labs(caption = "n = 855, span = .25")

p

ggsave("/Users/jcooper/Desktop/thesis_research/canopy_cover/2_canopy_traits/2019/results/2019_loess_CC.png", p, width=4, height=4)

```